---
title: "Initial Korean covid 19"
author: "Juliet Clark"
date: "5/30/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction



### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

$H_A$:Dying is dependent of sex when it comes to covid in Korea.

$H_0$:Dying is not dependent on sex when it comes to covid in Korea.

# Methods

result is a categorical variable, sex is categorical
result depends on sex'we use helper file cat~cat


## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")
```
This barchart shows that the frequency of women to men in comparison to being alive is greater, and being dead is lower than that of men. 

this barchart shows that...





  
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
this numerical summary shows that women were alive 12 more than men
## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The p-value for the chi-squared test is 0.001, so we reject the null hypothesis. 

The Fisher exact test shows that men are 2.3 times more likely to die than women in Korea.

## Conclusions
There is a relationship between sex and the amount of deaths due to Covid in Korea. 


